### 2023-02-13 chapters/background.tex
highlevel pass

A classifier $\alg$ that solely optimizes accuracy, i.e., the average number of times $\widehat{Y} = Y$, may discriminate certain compound sensitive groups over others~\cite{chouldechova2020snapshot}. In the following, we describe two well-known fairness definitions: group fairness and causal fairness. To this end, we use $ f(\alg, \mathcal{D}) $ to quantify the fairness of the classifier $ \alg $ given the distribution of features $ \mathcal{D} $. Alternatively, a fairness metric can be computed on a finite dataset instead of on the distribution. In that case, we use the notation $ f(\alg, \dataset) $, which is the bias of the classifier $ \alg $ on a dataset $ \dataset $.

@@

\paragraph{Disparate impact ($ \mathsf{DI} $).} Disparate impact also belongs to independence measuring group fairness metrics. Disparate impact measures the ratio between the minimum and the maximum probability of positive prediction of the classifier over all sensitive groups, and prescribe the ratio to be close to $1$~\cite{feldman2015certifying}.  Formally, the disparate impact of a classifier is 
\[
f_{\mathsf{DI}}(\alg, \mathcal{D}) \triangleq \frac{\min_{\mathbf{a}} \Pr[\widehat{Y} =1 | \mathbf{A} =  \mathbf{a}]}{\max_{\mathbf{a}} \Pr[\widehat{Y} =1 | \mathbf{A} =  \mathbf{a}]}.
\]

@@

\paragraph{Fairness Certification.} We certify the fairness of a classifier by comparing $ f(\alg, \mathcal{D}) $ with a fairness threshold, denoted by $ \epsilon \in [0,1] $, that quantifies the desired level of fairness. In particular, a classifier is $ \epsilon $-fair with respect to statistical parity, equalized odds, predictive parity, and path-specific causal fairness if and only if $ f(\alg, \mathcal{D}) \le \epsilon $. In contrast, a classifier achieves $(1 - \epsilon)$-disparate impact  if and only if $ f(\alg, \mathcal{D}) \ge 1 - \epsilon $. In all above fairness metrics, a lower value of $ \epsilon $ refers to higher fairness of the classifier.

>>>>

A classifier $\alg$ that solely optimizes accuracy, i.e. the average number of times $\widehat{Y} = Y$, may discriminate certain compound sensitive groups over others~\cite{chouldechova2020snapshot}. In the following, we describe two well-known fairness definitions: group fairness and causal fairness. To this end, we denote $ f(\alg, \mathcal{D}) $ to quantify the fairness of the classifier $ \alg $ given the distribution of features $ \mathcal{D} $. Alternatively, a fairness metric can be computed on a finite dataset instead of on the distribution. In that case, we use the notation $ f(\alg, \dataset) $, which is the bias of the classifier $ \alg $ on a dataset $ \dataset $.

@@

\paragraph{Disparate impact ($ \mathsf{DI} $).} Disparate impact also belongs to independence measuring group fairness metrics. Disparate impact measures the ratio between the minimum and the maximum conditional probability of positive prediction of the classifier over all sensitive groups, and prescribe the ratio to be close to $1$~\cite{feldman2015certifying}.  Formally, the disparate impact of a classifier is 
\[
f_{\mathsf{DI}}(\alg, \mathcal{D}) \triangleq \frac{\min_{\mathbf{a}} \Pr[\widehat{Y} =1 | \mathbf{A} =  \mathbf{a}]}{\max_{\mathbf{a}} \Pr[\widehat{Y} =1 | \mathbf{A} =  \mathbf{a}]}.
\]

@@

\paragraph{Fairness Verification.} We verify the fairness of a classifier by comparing $ f(\alg, \mathcal{D}) $ with a fairness threshold, denoted by $ \epsilon \in [0,1] $, that quantifies the desired level of fairness. In particular, a classifier is $ \epsilon $-fair with respect to statistical parity, equalized odds, predictive parity, and path-specific causal fairness if and only if $ f(\alg, \mathcal{D}) \le \epsilon $. In contrast, a classifier achieves $(1 - \epsilon)$-disparate impact  if and only if $ f(\alg, \mathcal{D}) \ge 1 - \epsilon $. In all above fairness metrics, a lower value of $ \epsilon $ refers to higher fairness of the classifier.

### 2023-02-13 chapters/fairness/fif/appendix.tex
highlevel pass

\appendix
\chapter{Fairness Influence Functions}
\section{Societal Impact}
In this paper, we quantify the influence of input features on the incurred bias/unfairness of the classifier on a dataset. This quantification facilitates our understanding of potential features or subsets of features attributing highly to the bias. This also allows us to understand the effect of the fairness enhancing algorithms in removing bias and their achievement/failure to do so. To the best of our understanding, this paper does not have any negative societal impact.

@@

\section{A Smoothing Operator $\textsc{Smooth}$: Cubic Splines}
\label{fairness_fairXplainer_sec:smoothing} 
In the $\textsc{LocalRegression}$ module of {\fairXplainer} (Line~\ref{fairness_fairXplainer_algo_line:local_regression_start}--\ref{fairness_fairXplainer_algo_line:local_regression_end}, Algorithm~\ref{fairness_fairXplainer_algo:framework}), we use a smoothing operator $\textsc{Smooth}$ (Line~\ref{fairness_fairXplainer_alg_line:backfitting_step}). In our experiments, \textit{we use cubic splines as the smoothing operator}. Here, we elucidate the technical details of cubic splines.

@@

\subsection{FIFs of Equalized Odds.}  For equalized odds, we deploy {\fairXplainer} twice, one for computing FIFs on a subset of data points in the dataset where $ Y = 1 $ and another on data points with $ Y = 0 $. Since, the maximum of the sum of FIFs between $ Y = 1 $ and $ Y = 0 $ is the equalized odds of the classifier, we finally report FIFs of features corresponding to the maximum sum of FIFs between $ Y \in \{0, 1\} $.

\subsection{FIFs of Predictive Parity.}  To compute FIFs for predictive parity, we condition the dataset by the predicted class $ \widehat{Y} $ and separate into two sub-datasets: $ \widehat{Y} = 1 $ and  $ \widehat{Y} = 0 $. For each sub-dataset, we deploy $ \fairXplainer $ by setting the ground-truth class $ Y $ as label. This contrasts the computation for  statistical parity and equalized odds, where the predicted class $ \widehat{Y} $ is considered as label. Finally, the maximum of the sum of FIFs between two sub-datasets for $ \widehat{Y} = 1 $ and  $ \widehat{Y} = 0 $ measures the predictive parity. Similar to equalized odds, FIFs achieving the greatest sum of FIFs for $ \widehat{Y} \in \{0, 1\} $ are the reported FIFs for the predictive parity of the classifier.

@@

	\caption[Ablation study on FIFs: $ \tau $]{Effect of spline intervals on the approximation error of statistical parity, root-mean square error (RMSE), and  execution time of {\fairXplainer}.}
	\label{fairness_fairXplainer_fig:effect_spline_intervals_appendix}
\end{figure}

@@

	\caption[Ablation study on FIFs: $ \lambda $]{Effect of maximum order $ \lambda $ on the approximation error of statistical parity, root-mean square error (RMSE) and  execution time of {\fairXplainer}.}
	\label{fairness_fairXplainer_fig:effect_maxorder_appendix}
\end{figure}

@@

\begin{figure}[!t]
	\begin{minipage}{0.45\textwidth}
		\centering
		\subfloat[Individual FIFs]{\includegraphics[scale=0.45]{figures/fairness/fif/adult_feature_weight_unsorted}\label{fairness_fairXplainer_fig:individual_fifs_adult}}
	\end{minipage}
	\begin{minipage}{0.5\textwidth}
		\centering
		\subfloat[Individual and intersectional FIFs]{\includegraphics[scale=0.45]{figures/fairness/fif/adult_feature_weight}\label{fairness_fairXplainer_fig:individual_and_intersectional_fifs_adult}}
	\end{minipage}
	\caption[FIFs for Adult dataset]{FIFs for Adult dataset on explaining statistical parity.}\label{fairness_fairXplainer_fig:individual_vs_intersectional_influence_adult}
\end{figure}

>>>>

\appendix
\chapter{Fairness Influence Functions}
%\section{Societal Impact}
%In this paper, we quantify the influence of input features on the incurred bias/unfairness of the classifier on a dataset. This quantification facilitates our understanding of potential features or subsets of features attributing highly to the bias. This also allows us to understand the effect of the fairness enhancing algorithms in removing bias and their achievement/failure to do so. To the best of our understanding, this paper does not have any negative societal impact.

@@

\section{A Smoothing Operator: Cubic Splines}
\label{fairness_fairXplainer_sec:smoothing} 
In the $\textsc{LocalRegression}$ module of {\fairXplainer} (Line~\ref{fairness_fairXplainer_algo_line:local_regression_start}--\ref{fairness_fairXplainer_algo_line:local_regression_end}, Algorithm~\ref{fairness_fairXplainer_algo:framework}), we use a smoothing operator $\textsc{Smooth}$ (Line~\ref{fairness_fairXplainer_alg_line:backfitting_step}). In our experiments, \textit{we use cubic splines as the smoothing operator}. Here, we elucidate the technical details of cubic splines.

@@

\subsection{FIFs of Equalized Odds}  For equalized odds, we deploy {\fairXplainer} twice, one for computing FIFs on a subset of data points in the dataset where $ Y = 1 $ and another on data points with $ Y = 0 $. Since, the maximum of the sum of FIFs between $ Y = 1 $ and $ Y = 0 $ is the equalized odds of the classifier, we finally report FIFs of features corresponding to the maximum sum of FIFs between $ Y \in \{0, 1\} $.

\subsection{FIFs of Predictive Parity}  To compute FIFs for predictive parity, we condition the dataset by the predicted class $ \widehat{Y} $ and separate into two sub-datasets: $ \widehat{Y} = 1 $ and  $ \widehat{Y} = 0 $. For each sub-dataset, we deploy $ \fairXplainer $ by setting the ground-truth class $ Y $ as label. This contrasts the computation for  statistical parity and equalized odds, where the predicted class $ \widehat{Y} $ is considered as label. Finally, the maximum of the sum of FIFs between two sub-datasets for $ \widehat{Y} = 1 $ and  $ \widehat{Y} = 0 $ measures the predictive parity. Similar to equalized odds, FIFs achieving the greatest sum of FIFs for $ \widehat{Y} \in \{0, 1\} $ are the reported FIFs for the predictive parity of the classifier.

@@

	\caption[Ablation study on FIFs: effect of spline intervals]{Effect of spline intervals on the approximation error of statistical parity, root-mean square error (RMSE), and  execution time of {\fairXplainer}.}
	\label{fairness_fairXplainer_fig:effect_spline_intervals_appendix}
\end{figure}

@@

	\caption[Ablation study on FIFs: effect of maximum order of intersectionality]{Effect of maximum order $ \lambda $ on the approximation error of statistical parity, root-mean square error (RMSE) and  execution time of {\fairXplainer}.}
	\label{fairness_fairXplainer_fig:effect_maxorder_appendix}
\end{figure}

@@

\begin{figure}
	\begin{minipage}{0.45\textwidth}
		\centering
		\subfloat[Individual FIFs]{\includegraphics[scale=0.45]{figures/fairness/fif/adult_feature_weight_unsorted}\label{fairness_fairXplainer_fig:individual_fifs_adult}}
	\end{minipage}
	\begin{minipage}{0.5\textwidth}
		\centering
		\subfloat[Individual and intersectional FIFs]{\includegraphics[scale=0.45]{figures/fairness/fif/adult_feature_weight}\label{fairness_fairXplainer_fig:individual_and_intersectional_fifs_adult}}
	\end{minipage}
	\caption[FIFs for Adult dataset]{FIFs for Adult dataset on explaining statistical parity.}\label{fairness_fairXplainer_fig:individual_vs_intersectional_influence_adult}
\end{figure}

### 2023-02-13 chapters/fairness/fif/fairxplainer.tex
introduction updated
highlevel pass

%\iffalse
\begin{algorithm}[t!]
	\caption{{\fairXplainer}}\label{fairness_fairXplainer_algo:framework}
	\begin{flushleft}
	\hspace*{\algorithmicindent}\textbf{Input:} Classifier $ \alg: (\nonsensitive, \sensitive) \rightarrow \widehat{Y}$, Dataset $\mathbf{D} = \{(\mathbf{z}^{(i)}, y^{(i)})\}_{i=1}^n$, Fairness metric $f(\alg, \mathbf{D}) \in \mathbb{R}^{\geq 0}$, Maximum order of intersectional influence $\lambda $\\
	\hspace*{\algorithmicindent}\textbf{Output:} FIF $ w_{\mathbf{S}}$ for the subsets of features $\{\feature_{\mathbf{S}}\}$
	\end{flushleft}

@@

		\Statex
		\Function{LocalRegression}{$ g_\mathbf{a}, \{\mathbf{z}^{(i)}\}_{i=1}^n, \lambda, m $}
		\label{fairness_fairXplainer_algo_line:local_regression_start}
		\State \textbf{Initialize}: $ \fiffunc_{\mathbf{a}, 0} \leftarrow \textsc{Mean}(\{g(\mathbf{z}^{(i)})\}_{i=1, \mathbf{a}^{(i)} = \mathbf{a}}^n), \widehat{\fiffunc}_{\mathbf{a}, \mathbf{S}} \leftarrow 0, \forall \mathbf{S} \in [m], \mathbf{S} \ne \emptyset, |\mathbf{S}|\le \lambda$ \label{fairness_fairXplainer_alg_line:initialization}
		\While{each $ \widehat{\fiffunc}_{\mathbf{a},\mathbf{S}} $ does not converge}
		\For{each $\mathbf{S}$}
		\State $ \widehat{\fiffunc}_{\mathbf{a}, \mathbf{S}} \leftarrow \textsc{Smooth}\big(\{\fiffunc_{\mathbf{a}}(\mathbf{z}^{(i)}) - \fiffunc_{\mathbf{a}, 0} - \sum_{\mathbf{S} \ne \mathbf{S}'} \widehat{\fiffunc}_{\mathbf{a}, \mathbf{S}'}(\mathbf{z}_{\mathbf{S}}^{(i)})\}_{i=1, \mathbf{a}^{(i)} = \mathbf{a}}^n\big)
		$\label{fairness_fairXplainer_alg_line:backfitting_step} \Comment{Backfitting}
		\State $\widehat{\fiffunc}_{\mathbf{a}, \mathbf{S}}  \leftarrow \widehat{\fiffunc}_{\mathbf{a}, \mathbf{S}} - \textsc{Mean}\big(\{\widehat{\fiffunc}_{\mathbf{a}, \mathbf{S}}(\mathbf{z}_{\mathbf{S}}^{(i)})\}_{i=1, \mathbf{a}^{(i)} = \mathbf{a}}^n\big)$ \label{fairness_fairXplainer_alg_line:mean_centered} \Comment{Mean centering}
		\EndFor
		\EndWhile
		\State \Return $ \fiffunc_{\mathbf{a}, \mathbf{S}},\fiffunc_{\mathbf{a}, 0} $
		\label{fairness_fairXplainer_algo_line:local_regression_end} 
		\EndFunction

		\Function{Covariance}{$\fiffunc_{\mathbf{a}},  \{\mathbf{z}^{(i)}\}_{i=1}^n, \fiffunc_{\mathbf{a}, \mathbf{S}},\fiffunc_{\mathbf{a}, 0}$}
		\label{fairness_fairXplainer_algo_line:covariance_computation_start}
		\State \Return $ \sum_{i=1, \mathbf{a}^{(i)} = \mathbf{a}}^n \fiffunc_{\mathbf{a}, \mathbf{S}}(\mathbf{z}_{\mathbf{S}}^{(i)})(g_\mathbf{a}(\mathbf{z}^{(i)}) - \fiffunc_{\mathbf{a}, 0}) $
		\label{fairness_fairXplainer_algo_line:covariance_computation_end}
		%		 \Comment{Eq.~\eqref{fairness_fairXplainer_eq:covariance_computation}}
		\EndFunction
	\end{algorithmic}
\end{algorithm}
%\fi
Here, $ \fiffunc_{\mathbf{a}, 0} $ is a constant, $ \fiffunc_{\mathbf{a},\mathbf{S}} $ is a \emph{component function} of $ \fiffunc_{\mathbf{a}} $ taking  a non-empty subset of features $ \feature_{\mathbf{S}} $ as input, and $ \delta $ is the approximation error. For computational tractability, we consider only components of \emph{maximum order} $ \lambda $, denoting the maximum order of intersectionality. {\fairXplainer} deploys backfitting algorithm for learning component functions in Eq.~\eqref{fairness_fairXplainer_eq:set_additive_classifier}, as discussed in the following.

>>>>

%\iffalse
\begin{algorithm}[t!]
	\small{
	\caption{{\fairXplainer}: An algorithm for estimating FIFs}\label{fairness_fairXplainer_algo:framework}
	\begin{flushleft}
	\hspace*{\algorithmicindent}\textbf{Input:} Classifier $ \alg: (\nonsensitive, \sensitive) \rightarrow \widehat{Y}$, dataset $\mathbf{D} = \{(\mathbf{z}^{(i)}, y^{(i)})\}_{i=1}^n$, fairness metric $f(\alg, \mathbf{D}) \in \mathbb{R}^{\geq 0}$, and maximum order of intersectional influence $\lambda $\\
	\hspace*{\algorithmicindent}\textbf{Output:} FIF $ w_{\mathbf{S}}$ for the subsets of features $\{\feature_{\mathbf{S}}\}$
	\end{flushleft}

@@

		\Statex
		\Function{LocalRegression}{$ g_\mathbf{a}, \{\mathbf{z}^{(i)}\}_{i=1}^n, \lambda, m $}
		\label{fairness_fairXplainer_algo_line:local_regression_start}
		\State \textbf{Initialize}: $ \fiffunc_{\mathbf{a}, 0} \leftarrow \textsc{Mean}(\{g(\mathbf{z}^{(i)})\}_{i=1, \mathbf{a}^{(i)} = \mathbf{a}}^n), \widehat{\fiffunc}_{\mathbf{a}, \mathbf{S}} \leftarrow 0, \forall \mathbf{S} \in [m], \mathbf{S} \ne \emptyset, |\mathbf{S}|\le \lambda$ \label{fairness_fairXplainer_alg_line:initialization}
		\While{each $ \widehat{\fiffunc}_{\mathbf{a},\mathbf{S}} $ does not converge}
		\For{each $\mathbf{S}$}
		\State $ \widehat{\fiffunc}_{\mathbf{a}, \mathbf{S}} \leftarrow \textsc{Smooth}\big(\{\fiffunc_{\mathbf{a}}(\mathbf{z}^{(i)}) - \fiffunc_{\mathbf{a}, 0} - \sum_{\mathbf{S} \ne \mathbf{S}'} \widehat{\fiffunc}_{\mathbf{a}, \mathbf{S}'}(\mathbf{z}_{\mathbf{S}}^{(i)})\}_{i=1, \mathbf{a}^{(i)} = \mathbf{a}}^n\big)
		$\label{fairness_fairXplainer_alg_line:backfitting_step}
		\State $\widehat{\fiffunc}_{\mathbf{a}, \mathbf{S}}  \leftarrow \widehat{\fiffunc}_{\mathbf{a}, \mathbf{S}} - \textsc{Mean}\big(\{\widehat{\fiffunc}_{\mathbf{a}, \mathbf{S}}(\mathbf{z}_{\mathbf{S}}^{(i)})\}_{i=1, \mathbf{a}^{(i)} = \mathbf{a}}^n\big)$ \label{fairness_fairXplainer_alg_line:mean_centered} \Comment{Mean centering}
		\EndFor
		\EndWhile
		\State \Return $ \fiffunc_{\mathbf{a}, \mathbf{S}},\fiffunc_{\mathbf{a}, 0} $
		\label{fairness_fairXplainer_algo_line:local_regression_end} 
		\EndFunction

		\Function{Covariance}{$\fiffunc_{\mathbf{a}},  \{\mathbf{z}^{(i)}\}_{i=1}^n, \fiffunc_{\mathbf{a}, \mathbf{S}},\fiffunc_{\mathbf{a}, 0}$}
		\label{fairness_fairXplainer_algo_line:covariance_computation_start}
		\State \Return $ \sum_{i=1, \mathbf{a}^{(i)} = \mathbf{a}}^n \fiffunc_{\mathbf{a}, \mathbf{S}}(\mathbf{z}_{\mathbf{S}}^{(i)})(g_\mathbf{a}(\mathbf{z}^{(i)}) - \fiffunc_{\mathbf{a}, 0}) $
		\label{fairness_fairXplainer_algo_line:covariance_computation_end}
		%		 \Comment{Eq.~\eqref{fairness_fairXplainer_eq:covariance_computation}}
		\EndFunction
	\end{algorithmic}
	}
\end{algorithm}
%\fi
Here, $ \fiffunc_{\mathbf{a}, 0} $ is a constant, $ \fiffunc_{\mathbf{a},\mathbf{S}} $ is a \emph{component function} of $ \fiffunc_{\mathbf{a}} $ taking  a non-empty subset of features $ \feature_{\mathbf{S}} $ as input, and $ \delta $ is the approximation error. For computational tractability, we consider only components of \emph{maximum order} $ \lambda $, denoting the maximum order of intersectionality. {\fairXplainer} deploys backfitting algorithm for learning component functions in Eq.~\eqref{fairness_fairXplainer_eq:set_additive_classifier}, as discussed in the following.

### 2023-02-13 chapters/fairness/justicia/experiments.tex
highlevel pass

\subsubsection{Verification: Detecting Compounded Discrimination in Sensitive Groups.}
We have tested {\justicia} for datasets consisting of multiple sensitive features and reported results in Figure~\ref{fairness_justicia_fig:sensitive_groups}. {\justicia} operates on datasets with even 40 compound sensitive groups and can potentially scale more than that while the state-of-the-art fairness verifiers (e.g., FairSquare and VeriFair) consider a single sensitive feature with two sensitive groups. Thus, {\justicia} removes an important limitation in practical fairness verification, which was previously restricted to Boolean sensitive groups. Additionally, in most datasets, we observe that disparate impact decreases and thus, discrimination increases as more compound sensitive groups are considered. For instance, when we increase the total  groups from $ 5 $ to $ 40 $ in the Adult dataset, disparate impact decreases from around $ 0.9 $ to $ 0.3 $, thereby detecting higher discrimination. Thus, {\justicia} detects that the marginalized individuals of a specific type (e.g., `race')  are even more discriminated and marginalized when they also belong to a marginalized group of another type (e.g., `sex').

@@

\subsubsection{Robustness: Stability to Sample Size.} 
We have compared the robustness of our probabilistic fairness verifier {\justicia} with dataset-centric verifier AIF360 by varying the sample-size and reporting the standard deviation of different fairness metrics. 
In Figure~\ref{fairness_justicia_fig:sample-size}, AIF360 shows higher standard deviation for lower sample-size and the value decreases as  the sample-size increases. 
In contrast, {\justicia} shows significantly lower ($\sim10\times$ to $100\times$) standard deviation for different sample-sizes. 
The reason is that AIF360 empirically measures on a fixed dataset whereas {\justicia} provides estimates over the distribution.
Thus, {\justicia} is more robust than the sample-based verifier AIF360.

@@

\subsubsection{Comparative Evaluation of Different Encodings.}
Both {\justiciaenum} and {\justicialearn}  have the same output according to Lemma~\ref{fairness_justicia_lm:equivalence}. However, {\justicialearn}  improves exponentially  in runtime  than {\justiciaenum} on both decision tree and Boolean CNF classifiers as we vary the total compound groups in Figure~\ref{fairness_justicia_fig:runtime_diff_encodings}. {\justiciacond} ({\justiciaenum} encoding where we consider conditional probabilities of non-sensitive features w.r.t. sensitive groups) also has an exponential trend in runtime similar to {\justiciaenum}.  This analysis justifies that the na\"ive enumeration-based approach cannot verify large-scale fairness problems containing multiple sensitive features, and {\justicialearn} is a more efficient approach for practical use.

>>>>

\subsubsection{Verification: Detecting Compounded Discrimination in Sensitive Groups}
We have tested {\justicia} for datasets consisting of multiple sensitive features and reported results in Figure~\ref{fairness_justicia_fig:sensitive_groups}. {\justicia} operates on datasets with even 40 compound sensitive groups and can potentially scale more than that while the state-of-the-art fairness verifiers (e.g., FairSquare and VeriFair) consider a single sensitive feature with two sensitive groups. Thus, {\justicia} removes an important limitation in practical fairness verification, which was previously restricted to Boolean sensitive groups. Additionally, in most datasets, we observe that disparate impact decreases and thus, discrimination increases as more compound sensitive groups are considered. For instance, when we increase the total  groups from $ 5 $ to $ 40 $ in the Adult dataset, disparate impact decreases from around $ 0.9 $ to $ 0.3 $, thereby detecting higher discrimination. Thus, {\justicia} detects that the marginalized individuals of a specific type (e.g., `race')  are even more discriminated and marginalized when they also belong to a marginalized group of another type (e.g., `sex').

@@

\subsubsection{Robustness: Stability to Sample Size} 
We have compared the robustness of our probabilistic fairness verifier {\justicia} with dataset-centric verifier AIF360 by varying the sample-size and reporting the standard deviation of different fairness metrics. 
In Figure~\ref{fairness_justicia_fig:sample-size}, AIF360 shows higher standard deviation for lower sample-size and the value decreases as  the sample-size increases. 
In contrast, {\justicia} shows significantly lower ($\sim10\times$ to $100\times$) standard deviation for different sample-sizes. 
The reason is that AIF360 empirically measures on a fixed dataset whereas {\justicia} provides estimates over the distribution.
Thus, {\justicia} is more robust than the sample-based verifier AIF360.

@@

\subsubsection{Comparative Evaluation of Different Encodings}
Both {\justiciaenum} and {\justicialearn}  have the same output according to Lemma~\ref{fairness_justicia_lm:equivalence}. However, {\justicialearn}  improves exponentially  in runtime  than {\justiciaenum} on both decision tree and Boolean CNF classifiers as we vary the total compound groups in Figure~\ref{fairness_justicia_fig:runtime_diff_encodings}. {\justiciacond} ({\justiciaenum} encoding where we consider conditional probabilities of non-sensitive features w.r.t. sensitive groups) also has an exponential trend in runtime similar to {\justiciaenum}.  This analysis justifies that the na\"ive enumeration-based approach cannot verify large-scale fairness problems containing multiple sensitive features, and {\justicialearn} is a more efficient approach for practical use.
